\documentclass{article}
\title{CSC352 HW7}
\author{Alex Zhang}
\date{March 2023}
\textwidth=16.00cm 
\textheight=22.00cm 
\topmargin=0.00cm
\oddsidemargin=0.00cm 
\evensidemargin=0.00cm 
\headheight=0cm 
\headsep=0.5cm
\textheight=610pt

\usepackage{latexsym,array,delarray,amsthm,amssymb,epsfig}
\usepackage{amsmath}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\newcommand{\bmat}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\mat}[1]{\mathbf{#1}}

\let\ds\displaystyle

\begin{document}
\maketitle

\section*{Question 1}
    \begin{enumerate}
        \item The relative error for $\mat{Q}$ is $2$.
        \item The relative error for $\mat{R}$ is $1.0014$.
        \item The relative error for $\mat{Q*R}$ is $1.2642e-15$.
    \end{enumerate}

    I'm surprised with the first two relative errors. For $\mat{Q}$'s relative error, it should be smaller than 1,
    but I got $2$, which shows there is a difference between true $\mat{Q}$, and calculated $\mat{Q}$. For $\mat{R}$, 
    I think it still should be $0$ for $\| 0\|_p = 0$. Base on the two relative errors, I think for HouseHolder QR, $\mat{Q}$ and $\mat{R}$ are not accurate.
    However, the relative error for $\mat{Q*R}$ is really small so their product is accurate. Based on this
    small relative error. This also says that Householder QR is stable.

\section*{Question 2}
    \begin{enumerate}
        \item For QR factorization with HouseHolder, the distance is $1.2316e-15$.
        \item For QR factorization with modified Gram-Schmidt, the distance is $0.2283$.
    \end{enumerate}

    The distance using HouseHolder is very small and therefore reasonable. However, the result for using mgs is quiet big.
    One reason is when doing Gram-Schmidt process, calculating matrix $\mat{Q}$ involves multiplication and normalization.
    This process will make $\mat{Q}$ not be strictly orthonormal matrix, and the result will be affected then. 

    Changing the value $\eta$, the distance from Gram and identity computed by mgs will decrease as $\eta$ increases.
    The distance of from Gram and identity through Householder will vary but only between $10^{-15}$ and $10^{-16}$.

\section*{Question 3}
    
Given that $\mat{L}$ and $\mat{B}$ are $\in \mathbb{R}^{n \times n}$, we write
a pseudocode function called LB:
\begin{lstlisting}
    function X = LB(L,B)
        for i = 1: length(B)
            for j = 1: length(B) do
                x(i,j) = (B(i,j)-$\sum^{j-1}_{k=1}$L(j,k)*x(i,k))/L(j,j)
            end
        end
    end fucntion
\end{lstlisting}
It eccentially adds a outer loop that helps store calculated $\mat{x}$ vector in each column. 
\end{document}